---
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Probability and Statistics

# Lab Assignment 1: Naive Bayes Classifier

### *Name1 Surname1, Name2 Surname2, Name3 Surname3*

## Introduction

During the past three weeks, you learned a couple of essential notions
and theorems, and one of the most important among them is the *Bayes
theorem*.

One of its applications is **Naive Bayes classifier**, which is a
probabilistic classifier whose aim is to determine which class some
observation probably belongs to by using the Bayes formula:
$$\mathsf{P}(\mathrm{class}\mid \mathrm{observation})=\frac{\mathsf{P}(\mathrm{observation}\mid\mathrm{class})\mathsf{P}(\mathrm{class})}{\mathsf{P}(\mathrm{observation})}$$

Under the strong independence assumption, one can calculate
$\mathsf{P}(\mathrm{observation} \mid \mathrm{class})$ as
$$\mathsf{P}(\mathrm{observation}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i), \qquad \mathsf{P}(\mathrm{observation} \mid \mathrm{class}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i \mid \mathrm{class}),$$
where $n$ is the total number of features describing a given
observation. Thus, $\mathsf{P}(\mathrm{class}|\mathrm{observation})$ now
can be calculated as

$$\mathsf{P}(\mathrm{class} \mid \mathrm{\mathrm{observation}}) = \mathsf{P}(\mathrm{class})\times \prod_{i=1}^{n}\frac{\mathsf{P}(\mathrm{feature}_i\mid \mathrm{class})}{\mathsf{P}(\mathrm{feature}_i)}\tag{1}$$

All the terms on the right-hand side can be estimated from the data as
respective relative frequencies;\
see [this
site](https://monkeylearn.com/blog/practical-explanation-naive-bayes-classifier/)
for more detailed explanations.

## Data description

There are 5 datasets uploaded on the cms.

To determine your variant, take your team number from the list of teams
on cms and take *mod 5* - this is the number of your data set.

-   **0 - authors** This data set consists of citations of three famous
    writers: Edgar Alan Poe, Mary Wollstonecraft Shelley and HP
    Lovecraft. The task with this data set is to classify a piece of
    text with the author who was more likely to write it.

-   **1 - discrimination** This data set consists of tweets that have
    discriminatory (sexism or racism) messages or of tweets that are of
    neutral mood. The task is to determine whether a given tweet has
    discriminatory mood or does not.

-   **2 - fake news** This data set contains data of American news: a
    headline and an abstract of the article. Each piece of news is
    classified as fake or credible. The task is to classify the news
    from test.csv as credible or fake.

-   **3 - sentiment** All the text messages contained in this data set
    are labeled with three sentiments: positive, neutral or negative.
    The task is to classify some text message as the one of positive
    mood, negative or neutral.

-   **4 - spam** This last data set contains SMS messages classified as
    spam or non-spam (ham in the data set). The task is to determine
    whether a given message is spam or non-spam.

Each data set consists of two files: *train.csv* and *test.csv*. The
first one you will need find the probabilities distributions for each of
the features, while the second one is needed for checking how well your
classifier works.

```{r}
# here goes a list of recommended libraries,
# though you may install other ones if they are needed
library(tidytext)
library(readr)
library(dplyr)
library(ggplot2)
library(tm)
library(wordcloud)
library(RColorBrewer)
library(wordcloud2)
install.packages("magrittr")
install.packages("dplyr")
library(magrittr)
library(dplyr)
```

## Instructions

-   The first step is data pre-processing, which includes removing
    punctuation marks and stop words

-   represent each message as a bag-of-words

-   using the training set, calculate all the conditional probabilities
    in formula (1)

-   use those to predict classes for messages in the test set

-   evaluate effectiveness of the classifier by calculating the
    corresponding metrics

-   shortly summarize your work

-   do not forget to submit both the (compiled) Rmd source file and the .html
    output
    
### Data pre-processing

-   Read the *.csv* data files.
-   Сlear your data from punctuation or other unneeded symbols.
-   Clear you data from stop words. You don't want words as is, and, or
    etc. to affect your probabilities distributions, so it is a wise
    decision to get rid of them. Find list of stop words in the cms
    under the lab task.
-   Represent each test message as its bag-of-words. Here:
    <https://machinelearningmastery.com/gentle-introduction-bag-words-model/>
    you can find general introduction to the bag-of-words model and
    examples on to create it.
-   It is highly recommended to get familiar with R dataframes, it would
    make the work much easier to do.
-   Useful links:
    -   <https://steviep42.github.io/webscraping/book/bagofwords.html#tidytext> -
        example of using *tidytext* to count frequencies of the words.
    -   Basics of Text Mining in R:
        <http://rstudio-pubs-static.s3.amazonaws.com/256588_57b585da6c054349825cba46685d8464.html>
        . Note that it also includes an example on how to create a bag
        of words from your text document.

```{r}
list.files(getwd())
list.files("data/0-authors")
```

```{r}
test_path <- "test.csv"
train_path <- "train.csv"

stop_words <- read_file("stop_words.txt")
# https://stackoverflow.com/questions/27195912/why-does-strsplit-return-a-list
splitted_stop_words <- strsplit(stop_words, split='\n')
splitted_stop_words <- splitted_stop_words[[1]]
```

```{r}
train <-  read.csv(file = train_path, stringsAsFactors = FALSE)
test <-  read.csv(file = test_path, stringsAsFactors = FALSE)
```

```{r}
# note the power functional features of R bring us! 
tidy_text <- unnest_tokens(train, 'splitted', 'Body', token="words")
#print(tidy_text)

tidy_text <- tidy_text %>% filter(!splitted %in% splitted_stop_words)
# now we have false+true for every meaningful word; need to split it to f and t

#print(tidy_text)

true_tidy <- tidy_text %>% filter(Label == "credible")
print(true_tidy)

false_tidy <- tidy_text %>% filter(Label == "fake")
print(false_tidy)

#print(splitted_stop_words)

#tidy_text %>% count(splitted, sort=TRUE)
true_tidy <- true_tidy %>% count(splitted, sort=TRUE)
colnames(true_tidy)[1] <- 'splitted'
colnames(true_tidy)[2] <- 'credible_num'
print(true_tidy)


false_tidy <- false_tidy %>% count(splitted, sort=TRUE)
colnames(true_tidy)[1] <- 'splitted'
colnames(true_tidy)[2] <- 'fake_num'
print(false_tidy)

total <- merge(true_tidy,false_tidy,by="splitted")
print(total)

```

### Data visualization

Each time you work with some data, you need to understand it before you
start processing it. R has very powerful tools to make nice plots and
visualization. Show what are the most common words for negative and
positive examples as a histogram, word cloud etc. Be creative!

## Classifier implementation




```{r}
naiveBayes <- setRefClass("naiveBayes",
                          
       # here it would be wise to have some vars to store intermediate result
       # frequency dict etc. Though pay attention to bag of wards! 
       fields = list(),
       methods = list(
                    # prepare your training data as X - bag of words for each of your
                    # messages and corresponding label for the message encoded as 0 or 1 
                    # (binary classification task)
                    fit = function(data)
                    {
                         data_fake <- train[train$Label == "fake",]
                         data_true <- train[train$Label == "credible",]
                         tidy_text_fake <- unnest_tokens(data_fake, 'splitted', 'Body', token="words") %>% filter(!splitted %in% splitted_stop_words)
                         tidy_text_true <- unnest_tokens(data_true, 'splitted', 'Body', token="words") %>% filter(!splitted %in% splitted_stop_words)
                         fake <- tidy_text_fake %>% count(splitted,sort=TRUE)
                         true <- tidy_text_true %>% count(splitted,sort=TRUE)
                         new_data <- fake %>% full_join(true, by="splitted")
                         new_data[is.na(new_data)] <- 0
                         num_col = nrow(new_data)
                         fake_sum <- sum(new_data$n.x)
                         true_sum <- sum(new_data$n.y)
                         new_data$prob_fake <- (new_data$n.x+1)/(num_col+fake_sum)
                         new_data$prob_true <- (new_data$n.y+1)/(num_col+true_sum)
                         
                         print(new_data)
                         return(new_data)
                    },
                    
                    # return prediction for a single message 
              predict = function(message)
                    {
                      # clear and split message
                      message <- gsub("[`[:punct:]]", '', message)
                      message <- tolower(message)
                      words = strsplit(message, " ")[[1]]
                      words <- words[!words %in% splitted_stop_words]
                      
                      return(words)
                      
                      # get constant numbers
                      total_fake_prob = total_number_of_fakes/total_number_of_new
                      total_cred_prob = 1 - total_fake_prob
                      
                      fake_stabilizer = sum(alldata$fakeNum) + nrows(alldata)
                      cred_stabilizer = sum(alldata$credNum) + nrows(alldata)
                      
                      credToFakeRatio = total_number_of_fakes/total_number_of_creds
                      
                      for (word in words){
                        
                        if ((word %in% alldata$splitted) & (alldata$fakeNum[alldata&splitted== word]>0))
                          credToFakeRatio = credToFakeRatio / data$fakeProb[data&splitted == word]
                        else 
                          credToFakeRatio = credToFakeRatio * fake_stabilizer
                        
                        if ((word %in% alldata$splitted) & (alldata$credNum[alldata&splitted== word]>0))
                          credToFakeRatio = credToFakeRatio * data$credProb[which(data&splitted == word)]
                        else
                          credToFakeRatio = credToFakeRatio / cred_stabilizer
                        
                      }
                      if (credToFakeRatio > 1){
                        return(1)
                        } else { return(0) }
                      },
                    
                    # score you test set so to get the understanding how well you model
                    # works.
                    # look at f1 score or precision and recall
                    # visualize them
                    # try how well your model generalizes to real world data! 
                    score = function(X_test, y_test)
                    {
                         # TODO
                    }
))

model = naiveBayes()
#model$fit()
fit_dataset <- model$fit()
```

```{r}
# score function
# we get some dataframe (X_test, y_test)

X_test_dataset <- head(tidy_text,30)[,2, drop=FALSE]
y_test_dataset <- head(tidy_text,30)[,3, drop=FALSE]

library(stringr)
test_function <- function(message) {
  #print(typeof(message))
  #print(message)
  #print("type")
  if(str_detect(message, "Trump Is So Busy")){
    #print(1)
    return (TRUE)
    
  }
  else{
    #print(0)
    return (FALSE)
    
  }
  
  #print(message)
  
  #return (nchar(message));
}

#merge two datasets into one

print(X_test_dataset)
print(y_test_dataset)

#print(typeof(X_test_dataset))

# uneffective
#n_of_rows = nrow(X_test_dataset)
#for (rowidx in 1:n_of_rows){
#  X_test_dataset[rowidx,1] <- lapply(X_test_dataset[rowidx,], test_function)
#}
#print(X_test_dataset)


#X_test_dataset['new_col'] <- NA
print(X_test_dataset)

print("testing..")
test_function("Trump 123")
print("tested")


X_test_dataset$predicted_rusults <- test_function(X_test_dataset$Headline)
print(X_test_dataset)

X_test_dataset$real = y_test_dataset$Label == "credible"
print(X_test_dataset)



print("test 2.1")
#X_test_dataset$real2.1 =
print("lapply..")
a <- lapply(X_test_dataset$Headline, test_function)  #str_detect("Trump", y_test_dataset$Label)
print(a)
print(typeof(a))

#X_test_dataset$real_final <- a
X_test <-cbind(w, id)
print(X_test_dataset)

df_results <- data.frame(a)
print(df_results)

X_test_dataset <- X_test_dataset %>% full_join(df_results)


print(X_test_dataset)



X_test_dataset$real2 = test_function(y_test_dataset$Label)
print(X_test_dataset)


print(X_test_dataset)



```


```{r}
#Create a vector containing only the text
#text <- data$text
# Create a corpus  
#docs <- Corpus(VectorSource(text))

test_path <- "test.csv"
train <-  read.csv(file = train_path, stringsAsFactors = FALSE)

library(wordcloud)
library(RColorBrewer)
library(wordcloud2)

# prepare data for visualization
data_fake <- train[train$Label == "fake",]
data_true <- train[train$Label == "credible",]
tidy_text_fake <- unnest_tokens(data_fake, 'splitted', 'Body', token="words") %>% filter(!splitted %in% splitted_stop_words)
tidy_text_true <- unnest_tokens(data_true, 'splitted', 'Body', token="words") %>% filter(!splitted %in% splitted_stop_words)
fake <- tidy_text_fake %>% count(splitted,sort=TRUE)
true <- tidy_text_true %>% count(splitted,sort=TRUE)
new_data <- fake %>% full_join(true, by="splitted")
new_data[is.na(new_data)] <- 0
num_col = nrow(new_data)
fake_sum <- sum(new_data$n.x)
true_sum <- sum(new_data$n.y)
new_data$prob_fake <- (new_data$n.x+1)/(num_col+fake_sum)
new_data$prob_true <- (new_data$n.y+1)/(num_col+true_sum)
fit_dataset <- new_data


print(fit_dataset)
graph_d <- fit_dataset
colnames(graph_d)[1] <- 'word'
graph_d$freq <- graph_d$n.x + graph_d$n.y
graph_d <- graph_d %>% select(-(n.x:prob_true))
print(graph_d)
# option 1
# - for all words we get
wordcloud2(graph_d, color = "random-light", backgroundColor = "grey", widgetsize = c(1000,1000))
# - only for fake words we get
graph_d$freq <- fit_dataset$n.x
print(graph_d)
wordcloud2(graph_d, color = "random-light", backgroundColor = "grey", widgetsize = c(1000,1000))
# - only for true words we get
graph_d$freq <- fit_dataset$n.y
print(graph_d)
wordcloud2(graph_d, color = "random-light", backgroundColor = "grey", widgetsize = c(1000,1000))


# option 2
# as an example, for fake words we get
set.seed(1000)
wordcloud(words = fit_dataset$splitted, freq = fit_dataset$n.x, min.freq = 20, max.words=250, random.order=FALSE, rot.per=0.15, colors=brewer.pal(8, "Dark2"))



```




## Measure effectiveness of your classifier
-   Note that accuracy is not always a good metric for your classifier.
    Look at precision and recall curves, F1 score metric.
-   Visualize them.
-   Show failure cases.

## Conclusions

Summarize your work by explaining in a few sentences the points listed
below.

-   Describe the method implemented in general. Show what are
    mathematical foundations you are basing your solution on.
-   List pros and cons of the method. This should include the
    limitations of your method, all the assumption you make about the
    nature of your data etc.
